---
title: "Problem writing Web Mercator"
author: "J. Beaulieu"
date: "5/6/2020"
output: html_document
---



## Problem preserving coordinate system with st_write

I'm working up some shapefiles in R.  The files will be written to disk, used to make a map in ArcGIS Pro, and the map will be published as a 'Web Layer' to the GeoPlatform.  Within the GeoPlatfrorm, I will use the Web Layer to create a Web Map.  ESRI advised that the shapefiles should use the Web Mercator projection (3857) to be consistent with the projection of the base maps.

Lets read in a .shp.  It comes into R with a 4267 EPSG code.

```{r }
library(sf)

nc <- st_read(system.file("shape/nc.shp", package="sf"))
st_crs(nc)
```

Lets project to 3857:

```{r }
nc <- nc %>% st_transform(3857)
st_crs(nc)

```

OK, good, now write to disk.
```{r}
st_write(nc, "nc.shp", delete_layer = TRUE)

```

When I read the file back into R, the coordinate system seems to be corrupted.  The EPSG code is gone and the coordinate system is no longer equal to 3857.  I first noticed this when working with the object in ArcGIS Pro which reported a coordinate system of Mercator 2SP.  It seems that st_write is somehow corrupting the coordinate system definition.  Any idea what's going on?
```{r}
nc.merc <- st_read("nc.shp")
st_crs(nc.merc)
st_crs(nc.merc) == st_crs(3857)
```




