---
title: "CRS with spSurvey V4.1.2"
author: "J. Beaulieu"
date: "4/1/2020"
output: html_document
---

## Load libraries and print session info
It is strange that R reports "Version 4.1.0 of the spsurvey package was loaded successfully." but sessionInfo() indicates that V4.1.2 is attached.  Is this a minor bug in the current CRAN package?
```{r}
library(spsurvey)
sessionInfo()
```

## Read the sf object

Read Utah ecoregions sf object that came with package.
```{r}
data("UT_ecoregions")
class(UT_ecoregions)
st_crs(UT_ecoregions) # EPSG 5070, CONUS Albers
```

## Attribute data
```{r}
head(UT_ecoregions)
```

## Unstratified, equal probability, GRTS survey design

Create design list:
```{r}
Equaldsgn <- list(None=list(panel=c(PanelOne=50), seltype="Equal"))
```


Select the sample.  This writes out a .shp and creates a SpatialDesign object.
```{r warning=FALSE}
Equalsites <- grts(design=Equaldsgn,
                   DesignID="EQUAL",
                   type.frame="area",
                   src.frame="sf.object",
                   sf.object=UT_ecoregions,
                   maxlev = 5,
                   shapefile=TRUE)
```

Print the initial six lines of the survey design:
```{r}
head(Equalsites)
```

Can convert SpatialDesign object to sf object, but sf object has no coordinate reference system.
```{r}
Equalsite.sf <- st_as_sf(Equalsites)

st_crs(Equalsite.sf)
```


The written .shp does not have a defined CRS either?
```{r}
foo <- st_read(dsn = ".", layer = "sample")
st_crs(foo)
```

Previous version of spSurvey had a prjfilename argument to specify projection of written .shp, but this is deprecated in 4.1.1.  How should I define CRS of design object?


