---
title: 'National Reservoir Methane Survey: R10'
author: "J. Beaulieu"
date: "12/16/2019"
output:
  html_document:
    df_print: paged
    toc: FALSE
    code_folding: hide
  editor_options: 
  chunk_output_type: console
---


```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(sf) # spatial data
library(tidyverse) # dplyr, ggplot
library(USAboundaries)

# Load data, from Olsen
dsn <- read_sf("../..//output/surveyDsgnOutput/SuRGE_design_20191206.shp")
```


ORD and Region 10 have previously measured CH4 and CO2 emission rates from reservoirs in Washington, Idaho, and Oregon.  The methods used during the surveys are identical to those that will be used for the new national survey.  Any previously sampled sites that happen to be included in the new national survey will not be resampled.

## Region 10 sites
The new survey design contains `r dsn %>% as.data.frame() %>% filter(PSTL_CODE %in% c("WA", "OR", "ID") & panel == "Base") %>% nrow()` Base sites in R10.
Two of these sites were sampled by R10 in 2018.  Both sites are in the Western Mountains ecoregion and will not be resampled during this new survey.  


``` {r}
# SITE_ID and lake names of R10 surveyed reservoirs.
r10 <- c("NLA17_WA-10038", "Baker Lake",
         "NLA17_WA-10006", "Swofford Pond",
         "NLA17_WA-10002", "Wapato Lake",
         "NLA17_OR-10006", "Phillips Reservoir",
         "NLA17_ID-10175", "Mormon Reservoir",
         "NLA17_ID-10032", "Lower Goose Reservoir",
         "NLA17_ID-10007", "Springfield Reservoir")

r10 <- data.frame(SITE_ID = r10[seq(1, 13, 2)],
                  R10_NAME = r10[seq(2, 14, 2)], 
                  stringsAsFactors = FALSE)

dsn %>% as.tibble() %>% 
  filter(panel == "Base") %>% # only Base sites
  merge(., r10) %>% # bring in r10 sites
  select(SITE_ID, R10_NAME, panel, AG_ECO9) # show matches

```

This leaves 4 new sites to sample in R10.  
``` {r}
# New sites to sample in R10.
r10.new <- dsn %>% as.data.frame() %>% 
  filter(panel == "Base", # only Base sites
         PSTL_CODE %in% c("WA", "OR", "ID"), # only in R10
         !(SITE_ID %in% c("NLA17_OR-10006", "NLA17_WA-10006"))) %>% 
  select(SITE_ID, panel, AG_ECO9) # show matches

r10.new
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
# Get states map
states <- us_states() %>%
  dplyr::filter(name %in% c('Washington','Oregon', 'Idaho')) %>% # CONUS
  st_transform(5070)


ggplot() +
  geom_sf(data = states,  size = 0.1) +
  geom_sf(data = filter(dsn, SITE_ID %in% r10.new$SITE_ID), size = 1) +
  geom_label() +
  ggtitle("Sites in R10 to sample for national survey")


```
```



Four of the oversample sites in the Xeric ecoregion were sampled by R10 in 2018.

``` {r}
# oversample sites previously sampled by R10

dsn %>% as.data.frame() %>% 
  filter(panel == "OverSamp") %>% # only Base sites
  merge(., r10) %>% # bring in r10 sites
  select(SITE_ID, R10_NAME, panel, AG_ECO9) # show matches



```
Oversample sites must be used in the order in which they were designated by the survey design algorithm, as indicated by the siteID order, within each design category (stratum column in design file).  If no oversample is available within a category, then use next OverSamp site in category of choice.  
* XER_GT_6m_GT_7: no previously sampled R10 sites are designated as oversample sites in this category.  
* XER_GT_6m_LE_7: NLA17_ID-10032 (Lower Goose Reservoir) is third in the list.
                  NLA17_WA-10002 (Wapato Lake) is 9th in the list
                  NLA17_ID-10075 (Mormon Reservoir) is 14th in the list  
* XER_LE_6m_GT_7: no previously sampled R10 sites are designated as oversample sites in this category.  
* XER_LE_6m_LE_7: NLA17_ID-10007 (Springfield Reservoir) is 3rd in oversample list.

It is possible that previously sampled R10 sites will be used as oversample sites, most likely in XER_GT_6m_LE_7 and XER_LE_6m_LE_7.  


``` {r}
# Oversample sites in XERIC ecoregion.
dsn %>% as.tibble() %>% 
  filter(panel == "OverSamp", AG_ECO9 == "XER") %>%
  select(siteID, stratum, panel, SITE_ID, AG_ECO9) %>%
  print(n= Inf)

```

